---
title: Linear Regression for TriPASS Data Sci 20180925 Meetup
author: "Rick Pack"
date: "September 24, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
```

## Topic

Co-presentation with Kevin Feasel (linear regression). R code for regression of Carolina Godiva Summer Track Series (2018) data.

```{r trackres}
track_res <- read_csv('godiva_summer_track_res_2018.csv')
track_res %>% 
    # Differentiate walking events wit a bit of added distance
    mutate(dist_m = case_when(
                grepl("WALK", toupper(Event)) ~ as.integer(dist_m + 0.11),
                TRUE ~ dist_m)) %>%
    group_by(Name, Sex, Event, Date_Meet) %>% 
    summarise(ct = n()) %>% dplyr::filter(ct > 1)
# Example of name appearing errantly more than once for same event / day
track_res %>% dplyr::filter(str_trim(Name) == 'Brendan Murray' & str_trim(toupper(Event)) == '5K RUN')
track_res <- track_res %>% dplyr::filter(!is.na(Age)) %>%
    distinct(Name, Sex, Event, Date_Meet, .keep_all = TRUE)
track_res <- track_res %>% group_by(Name) %>% arrange(Name, Date_Meet, ct_evt) %>%
               mutate(EventNum = row_number()) %>% 
               group_by(Name, Date_Meet) %>%
               mutate(EventDay = row_number()) %>%
               group_by(Name, dist_m, Date_Meet) %>%
               mutate(DistNum = row_number()) %>%
               ungroup()

all_dates <- sort(unique(track_res$Date_Meet))
weather_2018_ht <- read_html('https://www.carolinagodiva.org/index.php?page=track-season-weather-conditions')
temps_2018      <- weather_2018_ht %>% html_nodes(xpath="//tr[(((count(preceding-sibling::*) + 1) = 12) and parent::*)]//td[(((count(preceding-sibling::*) + 1) = 2) and parent::*)]//span | //tr[(((count(preceding-sibling::*) + 1) = 11) and parent::*)]//td[(((count(preceding-sibling::*) + 1) = 2) and parent::*)]//span | //tr[(((count(preceding-sibling::*) + 1) = 10) and parent::*)]//td[(((count(preceding-sibling::*) + 1) = 2) and parent::*)]//span | //tr[(((count(preceding-sibling::*) + 1) = 9) and parent::*)]//td[(((count(preceding-sibling::*) + 1) = 2) and parent::*)]//span | //tr[(((count(preceding-sibling::*) + 1) = 8) and parent::*)]//td[(((count(preceding-sibling::*) + 1) = 2) and parent::*)]//span | //tr[(((count(preceding-sibling::*) + 1) = 7) and parent::*)]//td[(((count(preceding-sibling::*) + 1) = 2) and parent::*)]//span | //tr[(((count(preceding-sibling::*) + 1) = 6) and parent::*)]//td[(((count(preceding-sibling::*) + 1) = 2) and parent::*)]//span | //tr[(((count(preceding-sibling::*) + 1) = 5) and parent::*)]//td[(((count(preceding-sibling::*) + 1) = 2) and parent::*)]//span | //tr[(((count(preceding-sibling::*) + 1) = 4) and parent::*)]//td[(((count(preceding-sibling::*) + 1) = 2) and parent::*)]//span | //tr[(((count(preceding-sibling::*) + 1) = 3) and parent::*)]//td[(((count(preceding-sibling::*) + 1) = 2) and parent::*)]//span") %>% 
    html_text() %>%
    as.numeric() %>%
    data.frame() %>%
    # Season averages of somehow been pulled in, so removing rows exceeding the count of 
    # Summer Track Series dates
    slice(1:length(all_dates))
colnames(temps_2018) <- 'Temp'
temps_2018$Date_Meet <- all_dates    
track_res <- left_join(track_res, temps_2018, by = 'Date_Meet')

track_res %>% dplyr::filter(Name=='Rick Pack' & Date_Meet==ymd('2018-08-01'))

track_res_base <- track_res %>% 
    mutate(Sex_num = case_when(
        Sex == "M" ~ 0,
        Sex == "F" ~ 1)) %>%
    select(Sex_num, Age, dist_m, track_time, DistNum, EventNum, EventDay, Temp)
track_lm <- lm(track_time ~ . , track_res_base)
summary(track_lm)
# What rows have missing data given lm reported 246 'observations deleted due to missingness'?
which(! complete.cases(track_res_base))
track_res_base[c(3, 20, 27), ]

track_res_base_complete <- data.frame(track_res_base[complete.cases(track_res_base),])
track_res_base_complete$Predicted <- fitted(track_lm)
head(track_res_base_complete %>% select(Age, dist_m, track_time, Predicted))
```

## Flawed model, let's look at plots (the usual first step)
P-value for intercept is significant, suggests model is flawed

```{r plot1}
plot(track_res_base)
```

### No variable looks linear to the response variable track_time (column 4) but the combination of explanatory parameters could be.
```{r plot2}
lattice::xyplot(track_time ~ Predicted, track_res_base_complete, auto.key = TRUE)
```

#### We need to pursue more adjustments, such as developing separate models for shorter events (lower track_time) vs. longer events, dropping insignificant predictors, and possibly a different kind of model than a linear one.